### 2019-02-27 docs/notebooks/results.Rmd

## Summary
Convolutional neural networks (CNNs) represent the state of the art in image recognition. Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015]. This may be surpising to scientists studying the visual system, who typically describe attention as a result of the limitated resources available for processing sensory inputs. These scientists have developed behavioral tasks to investigate visual attention. Such tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how proposed attentional mechanisms change the output of neural networks. Here we adapt a classic form of visual search task so it can be performed by neural networks. We measure the accuracy of two CNNs, AlexNet and VGG16, as they report whether a target is present among distractors. These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain when presented with the same visual stimuli. Surprisingly, we find that both networks show attention-like effects when performing the task.

>>>>

## Abstract
Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation. However, they suffer from known flaws. Many mechanisms proposed to improve the performance of neural networks have been given the name "attention" [@xuShowAttendTell2015]. This may be surpising to scientists studying the visual system, who have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs. Attention in the brain has been primarily investiaged using visual search tasks. These tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs. We are pursuing these alternate benchmarks for attentional mechanisms in the course of our investigations of continual learning algorithms. Here we adapt a classic form of visual search task so it can be performed by neural networks. We measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors. These architectures were chosen in part because they have been used in several studies that report similarities between activations in the hidden layers of the network and activity in the visual system of the brain, when both are presented with the same images. Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the task. When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases. These dependency of accuracy on the number of items is often referred to as "set size effects". We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects. It might seem obvious that target-distractor similarity would impair accuracy, but much of the visual attention literature boils down to a debate about the extent to which these effects can be attributed to cognitive processes, e.g. whether those processes are serial or parallel and noisy. Here, since we know that CNNs process their inputs in parallel, we can conclusively attribute the impaired accuracy here to target-distractor similarity. Therefore, we suggest that any mechanism which can be reformulated as a learned weighting processed in parallel is not categorically different from current CNNs. Such mechanisms may not ever improve performance of CNNs, given that the attentional effects they produce will be of the the kind we describe here.
